---
title: "Assessment 2"
author: "STAT5009"
date: "2024-05-02"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Q1. Write your interpretation (one sentence) regarding the quality of the regression model based on the above residual plots. 

Both plots of representation models appear to be good and the residuals are scattered randomly around the horizontal zero line suggesting that the errors are independent and normally distributed while both can be used as a key assumption for a liner regression model.


### Q2. 

## a) load the airquality data and store in your R-session as data.frame called'air'. Show the dimensions of ‘air’ 


```{r, echo=TRUE, eval=TRUE}

air <- data.frame(airquality)
dim(air)

```

## b) Remove all missing values from this data and store the data as 'air.no'. Show the code


```{r, echo=TRUE, eval=TRUE}

str(air)
head(air)
summary(air)

air.no <- na.omit(air)
summary(air.no)

```


### Q3. 
Compare two approaches - backward regression and the leave-one-out cross validation (LOOCV) - to recommend the best model to predict Ozone concentration using other predictors. 


```{r, echo=TRUE, eval=TRUE}

library(leaps)

air.bw <- regsubsets(data = air.no,
                     Ozone ~.,
           method = "backward"
           )

air.s <- summary(air.bw)

names(air.s)

plot(air.s$adjr2,
     xlab = "No. of variables",
     ylab = "Adjusted R-sq",
     pch = 16,
     cex.lab = 1.5  
     )

```


The plot created indicates that the R-sq keeps rising as we add more variables and the rate of growth slows beyond the fourth variable that already explains 60% of the variation. It is possible to select all 5 variables in a model. Yet the impact on the 4th and 5th variables are considerably lower, it is best to omit the last variable.



```{r, echo=TRUE, eval=TRUE}

air.feat <- coef(air.bw,4) 
names(air.feat)

```

```{r, echo=TRUE, eval=TRUE}

airbw.mod <- lm(Ozone ~ Solar.R  +  Wind  +  Temp + Month,
                data = air.no
                )

summary(airbw.mod)

```


From the summary it is clear that the month is not significant as the P value is more than 0.05(actual value - 0.051).
Hence creating a new linear model without the month named airbw.mod2.


```{r, echo=TRUE, eval=TRUE}

airbw.mod2 <- lm(Ozone ~ Solar.R  +  Wind  +  Temp,
                data = air.no
                )

summary(airbw.mod2)

```
multi-collinearity check for the linear model "airbw.mod2"

```{r, echo=TRUE, eval=TRUE}

library(car)

vif(airbw.mod2)

```

The variation inflation factor is less that 5 and closer to 1 being a smaller mult-colinearity. Hence it is possible to use all the variances. 

residual analysis - plot of residuals versus fitted values

```{r, echo=TRUE, eval=TRUE}

fit <- fitted.values(airbw.mod2)
res <- residuals(airbw.mod2)
plot(airbw.mod2)

```

From the Residuals versus Fitted values plot shown above it is clear that the model is not linear. It is possible to observe that this is in a funnel shape meaning, linear regression model does not follow a constant variance. 



Model equation and interpretation of chosen model

```{r, echo=TRUE, eval=TRUE}

summary(airbw.mod2)

```

Using the summary above, can select any one feature coefficient to write interpretation of its relationship with Ozone.

Ozone = -64.34 + 0.06* Solar.R + (-3.33)* Wind + 1.65* Temp

Selecting Solar.R coefficient,
            Ozone will increase by 0.06 units with increase of 1 unit of Solar.R
            
Confidence interval of chosen model,

```{r, echo=TRUE, eval=TRUE}

confint(airbw.mod2)

```
according to the information above, 95% of the confidence will lay between 0.014 and 0.106


Cross-validation based selection:
R-implementation code

Using first three variables only since the results from previous activities shows that the significance of the variable 4, 5, is unaccountable.

```{r, echo=TRUE, eval=TRUE}

library(boot)

err_j <- c()
for(j in 1:3)
  
{
  air.j <- air[,1:(j+1)]
  airmod_j <- glm(Ozone ~., data = air.no)
  cv.j <- cv.glm(airmod_j, data = air.no)
  err_j[j] <- cv.j$delta[1]
}

str(cv.j)

plot(1:3, err_j)
  
```
```{r, echo=TRUE, eval=TRUE}

mse.air <- mean(err_j)
mse.air

```
computed Mean Squared Error for the cross validation according to above is 462.9437


```{r, echo=TRUE, eval=TRUE}

res.val.reg <- airbw.mod2$residuals

sq.err <- res.val.reg^2

mse.reg <- mean(sq.err)
mse.reg

```
Computed Mean Squared Error for the backward regression according to above is 432.4576

Hence the MSE is less in backward regression.



### Q4. 

Using the IRIS data (available with baseR) a botanist wants to assess if a given iris flower is of the virginica species -based on its morphological features, Sepal Length, Sepal Width, Petal Length, and Petal Width. S/he hired a statistician who did some modelling (below). 


```{r, echo=TRUE, eval=TRUE}

ir <- iris

str(ir)

```

## A.	Construct a categorical variable ‘vir’ that takes values ‘1’ if a flower is Virginica, ‘0’ otherwise. 

```{r, echo=TRUE, eval=TRUE}

vir <- ifelse(ir$Species == "virginica", 1, 0)

```


## B.	Assuming that this data is a true representation of the natural world what is the probability that a randomly selected flower is from the species virginica.  Show your code and result.

```{r, echo=TRUE, eval=TRUE}

table(ir$Species)

tab.ir <- table(ir$Species)/nrow(ir)
tab.ir

```

The probability of a randomly selected flower to be from the specie virginica is 33%. The amount of virginica flowers divided by the amount of total flowers in general will give the same answer.



## Q4.1- Which of the above would be your recommended model? Write out the full expression of this model. For your recommended model choose any (one) feature and write the interpretation of the corresponding model coefficient in relation to virginica flower. 


The model 1 uses all the variables while having a statistically significant Petal.Length and a Residual deviance of 11.899 and an AIC (Akaike Information Criterion) of 21.899.
The model 2 uses data while omitting the Sepal.Length, and Petal.Length and petal.Width both are statistically significant. The Residual deviance is 13.266 and the AIC is 21.266.
It is possible to conclude that the model 2 is recommended as the Residual deviance stays less in the model two meaning less deviation, while having a higher AIC value.



## Q4.2– Use your recommended model to predict the odds of the following IRIS flower being Virginica. Hint: You would need to manually calculate the odds, using the R-model estimates.

log.odds = (-50.527) + (-8.376)* Sepal.Width + 7.875* Petal.Length + 21.43* Petal.Width

log.odds = (-50.527) – (8.376*2.77) + (7.875*4.26) + (1.33*21.43)

log.odds = (-50.527) + 33.5475 - 23.20152 + 28.5019

 log.odds= -11.67912
odds = exp(-11.67912)

odds = 8.468815e-06
